---
title: "Creating Stories From Data"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Once we are completed our data analysis and modeling activities, we need to present the same to the wider audiences. For working professionals, the audience can vary from internal team members to the leadership team. To engage with them, we need to build stories using the model findings and insights. And, one of the important elements of those stories is going to be visualisation.

In R, there are a number of packages which we can utilise to create compelling plots. We will primarily be focusing on **ggplot2** and **plotly** here.


## Introduction to ggplot2

ggplot2 is part of **tidyverse** family of packages created for Data Science. It helps us to create graphics based on the data provided to it. Here, mostly we will use examples from 'Data Visualisation' chapter of 'R for Data Science' book written by Hadley Wickham and Garrett Grolemund. 

### How does ggplot2 works?

Let us install the library first.

```{r eval=FALSE}
#install.packages("ggplot2")
library(ggplot2)
```

Let's take R built-in dataset *mpg* to understand how to create different charts/plots using the package.

```{r}
data("mpg")
str(mpg)
```

We have information about manufactures, models, classes and types of cars in this dataset.

```{r}
head(mpg, 5)
```
  
Let's say we want to have a scatter plot between the engine displacement (*displ* column) and highway miles (*hwy* column). To draw this scatter plot, let us see how **ggplot** works in a step-by-step manner.

```{r}
p <- ggplot(data = mpg,
            aes(x = displ, y = hwy))
p
```

Well, the above code has created an empty graph with x and y axis as *displ* and *hwy* respectively. 


#### What did we do here?
We created a ggplot object, specifying the data frame to use to be *mpg* and aesthetic mapping *aes*. The mapping is to mention how the variables in the data are mapped to visual properties.

#### What do we need to bring the scatter plot?
We need to add what do we want to be on the graph to the existing ggplot object. We will have to add *geom_point* function for creating scatter plot. *geom_*functions are the geometrical properties for bar, line, scatter plots in the ggplot system.

```{r}
p <- p + geom_point()
p
```


#### What if we wanted to have different colours for different classes?

We are going to plot the scatter plot between *displ* and *hwy*, but we want to identify the classes of the cars from the graph as well.

```{r}
p <- ggplot(data = mpg,
            aes(x = displ, y = hwy, colour = class)) +
  geom_point()
p
```

What have we done here? We have created another dimension *class* in the aesthetics of ggplot arguments. Now, we can draw some insights from *displ* vs *hwy* pattern on different classes of cars as well.


#### What if we wanted to have different shapes and colours for different classes?

Let us see what is the argument needs to be put in the aesthetics of ggplot.

```{r warning=FALSE}
p <- ggplot(data = mpg,
            aes(x = displ, y = hwy, shape = class, colour = class)) +
  geom_point()
p
```

We added both *shape* and *colour* arguments in the aesthetics and here's the plot. We notice, there is no shape assigned for the class *suv*. The reason is, ggplot2 will only use six shapes at a time. 

#### Alternate way to write the code

So far I have written the aesthetics inside the first part of the code snippet. What if we specify the aesthetic inside the second part, as the *geom* is actually clarifying which kind of plot we want for our data.

```{r}
p <- ggplot(data = mpg) +
  geom_point(aes(x = displ, y = hwy, colour = class))
p
```

The above command is also giving the same plot for *displ* vs *hwy*.

***

## Introduction to plotly

Another way we can create interactive graphics is using the package named **plotly**. This package uses the open-source JavaScript graphing library *plotly.js*.

### How does plotly works?

Let us install the library first.

```{r eval=FALSE}
#install.packages("plotly")
library(plotly)
```


Let us take the same *mpg* data to understand how to create charts using plotly. Let's say we want to craete the same scatter plot between the *displ* and *hwy*. 

```{r message=FALSE}
fig <- plot_ly(data = mpg,
            x = ~hwy, y = ~displ)
fig
```

#### What is different here?
If we hover over the points present on the scatterplot, we will see the (x,y) points. 

#### What if we wanted to have different colours for different classes?

We are going to write in a similar fashion like ggplot to see the scatter plot between *displ* and *hwy*, with identifying classes of the cars from the graph as well.

```{r message=FALSE}
fig <- plot_ly(data = mpg,
            x = ~hwy, y = ~displ, colour = ~class)
fig
```


## Data Usage

**coming soon** 